\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-14 trunk/arch.tex
updates

\begin{quotation}
  The planner is not concerned with any change that may occur in
  $\Sigma$ \footnote{the world modeled by the plan domain} while it is
  planning; it plans for the given initial and goal states regardless
  of the current dynamics, if any. 
\end{quotation}

@@

\subsubsection{Deliberation : planning for future state evolution}
\label{sec:arch:plan}

The planning process inside a reactor is directly based on the way
\eu work which was described in previous section. Still some simple
alterations were made in order to both include deliberation
information one reactor provide to the architecture -- such as its
deliberation latency and its planning look-ahead both expressed in
ticks -- and the fact that we potentially need to interleave this
deliberation with not the synchronization process but also with the
deliberation of other reactors\footnote{Current implementation of \rx
  is running on a single process and it is the responsibility of the
  agent itself to emulate the multi-threading of the reactors
  deliberation and synchronization}. For this reason while the overal
planning of a \eu reactor strictly relies on the \eu framework
plnanning solver, its slices the execution of Algorithm
\ref{alg:europa:solve} in atomic steps. The interruption is done
around the recursive call (line \ref{li:recurse}) while always
ensuring that the current partial plan mainitained is never proven
inconsistant at any single step. This can be managed by allowing the
call to backtrack until a DecisionPoInt that is not exhausted (ie
$decision.hasNext()$ is \texttt{true}) is find in the call stack.

One critical aspect of this deliberation step is to ensure that at the
exit of this call the $plan$ is not proven inconsistent. Indeed, it is
always possible that the next call made by the agent is the
synchronization which would immediately fail if the $plan$ is
inconsistent before any solving attempt made for synchronization.  For
this reason if the plan is proven inconsistent during a deliberation
step we either backtrack in the decision stack until we find a node fo
the tree that have remaining alternate solution or simply fully relax
the plan in situation there's absolutely no other alternative (by
removing all the goals and keeping only recent observations
produced). These constraints are yet again related to the fact that
for the agent the only crucial functionality of a reactor is to
resolve successfully its synchronization.

The number of flaws between two steps can evolve for the following
reasons that are exogenous to deliberation :

@@

\item A new goal has been posted on one of the reactor timeline. This
  is usually due to another reactor finding a plan and posting the
  corresponding new objective to this reactor.

@@

\item A previously requested goal has be recalled by the initial
  requester. This can happen when the reactor did identify that is
  initial plan is no longer valid. For example as synchronization
  proved this plan to be inconsistent.

@@

These 3 events will potentially produce new flaws to be resolved while
any single deliberation step will usually reduce the number of flaws
to be resolved.

The scope of planning for the reactor plays also a role in the
deliberation process as we limit deliberation to the sliding temporal
window $[\tau+\lambda, \tau+\lambda+\pi]$ where:

@@

\item $\tau$ is the current tick

\item $\lambda$ is the specified latency of this reactor. Which is an
  indicator on the expected maximum time this reactor is expected
  before being able to produce a plan.\footnote{Do note that this
    parameter remains indicative and, while it is highly recommended
    to select a sound value, a failure to produce a plan in this delay
    is not considered as a critical failure of the reactor}

\item $\pi$ is the planning look-ahead of the reactor. which indicates
  how many ticks in the future this reactor is looking ahead while planning.

@@

This scope allow to filter out any tokens that are either necessarily
ending before $\tau$ or necessarily starting after. Focusing the
[planning problem to only the tokens that are in a reasonable
future. This window is taken into account by the agent which will
notify the reactor of new {\em internal} goals only when they overlap
this window but also at the deliberation solver in order to reduce the
number of tokens to be evaluated and consequently reduce the cost of
deliberation. As time advance flaws that where initially beyond this
window will become active flaw that the reactor can then evaluate
resulting on an apparent continuous planning of the reactor.

When there are no more flaw is present in the current planning scope a
plan solution is found. This has two effects:

@@

\item The {\em external} part of this plan can then be posted which
  eventually (at the end of the current tick if these goals overlap
  the owner planning window) will be dispatched as goals triggering
  new deliberation on these reactors.

>>>>

\begin{quotation}
  The planner is not concerned with any change that may occur in
  $\Sigma$ \footnote{The world modeled by the plan domain.} while it is
  planning; it plans for the given initial and goal states regardless
  of the current dynamics, if any. 
\end{quotation}

@@

\subsubsection{Deliberation: Planning for future state evolution}
\label{sec:arch:plan}

Deliberation inside a reactor is based on \eu capabilities described
in Section \ref{sec:europa:arch}. However deliberation within \rx
comes with some augmentation -- deliberation information one reactor
provide to the architecture such as its deliberation latency and its
plan look-ahead both expressed in ticks -- and the fact that we
potentially need to interleave deliberation with not just the
synchronization process but also with deliberation of other
reactors\footnote{The current implementation of \rx is running on a
  single process and it is the responsibility of the agent itself to
  emulate reactor multi-threading for deliberation and
  synchronization.}. For this reason while the overall planning of a
deliberative reactor strictly relies on the \eu framework planning
solver, it slices the execution of Algorithm \ref{alg:europa:solve} in
atomic steps. The interruption is done around the recursive call (line
\ref{li:recurse}) while ensuring that the current partial plan is
never proven inconsistant at any single step. This can be managed by
allowing the call to backtrack until a \textsf{DecisionPoInt} that is
not exhausted (\ie $decision.hasNext()$ is \texttt{true}) is found in
the call stack.

A critical aspect of this deliberation step is to ensure that at the
exit of this call the $plan$ is not inconsistent. % Indeed, it is always
% possible that the next call made by the agent is for synchronization
% which could fail if the $plan$ is inconsistent before any attempt is
% made for synchronization.
If the partial plan is found to be inconsistent during a deliberation
step we either backtrack in the decision stack until a node of the
tree is found with an alternate solution or the plan is relaxed since
there is no other alternative solution. We do this by removing all
goals and keeping only recent observations. % These constraints are yet
% again related to the fact that for the agent the only crucial
% functionality of a reactor is to resolve successfully its
% synchronization.

The number of flaws between two steps \comment{between deliberation
  and synch.?} can evolve for the following reasons that are exogenous
to deliberation:

@@

\item A new goal has been posted on a reactor timeline. This is
  usually due to another reactor finding a partial plan and posting
  the corresponding new objective.

@@

\item A previously requested goal has be recalled by the initial
  requester. This can occur when the reactor's initial plan is no
  longer valid, for instance with synchronization resulting in
  inconsistency.

@@

\comment{this might be a subsubsection before someplace since it
  introduces the concept of latency/lookahead.}
These events could potentially produce new flaws to be resolved while
any single deliberation step will likely reduce the number of flaws
for resolution. The scope of planning for a reactor also plays a role
in the deliberation process as we limit deliberation to the sliding
temporal window $[\tau+\lambda, \tau+\lambda+\pi]$ \cite{py10} where:

@@

\item $\tau$ is the current tick indicative of the execution frontier

\item $\lambda$ is the specified latency of this reactor, which is an
  indicator of the maximum expected time for a reactor to produce a
  plan.\footnote{Note that while it is highly recommended to select a
    sound value, a failure to produce a plan within a chosen value for
    this parameter is not considered as a critical failure of the
    reactor.}

\item $\pi$ is the planning look-ahead of the reactor. which indicates
  how many ticks in the future this reactor is looking ahead during
  deliberation.

@@

Such a scope allows filtering of tokens that are either necessarily
ending before $\tau$ or necessarily starting after, focusing the
planning problem on only those tokens that are within a ``reasonable''
future. This window is taken into account by the \rx agent which will
notify a reactor of new {\em internal} goals not only when they
overlap this reactor specific window but also for the deliberation
solver in order to reduce the number of tokens to be evaluated to
reduce deliberation cost and performance. As time advances flaws that
were initially beyond this temporal horizon will become active which a
reactor can evaluate. This results in an apparent \emph{continuous
  planning} of the reactor. When there are no more flaws present in
the current planning scope a partial plan solution is found. This has
two effects:

@@

\item The token on the {\em external} timelines of this partial plan
  can then be posted which eventually (at the end of the current tick
  if these goals overlap the owner reactor's planning window) will be
  dispatched as goals, in turn triggering further deliberation.
